---
title: "Pondo_Allom_Params"
author: "Katie Murenbeeld"
date: "3/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:

## Install the BAAD Package:

Install the package from github:

```{r install, echo=TRUE, eval=TRUE}
#install.packages("devtools")
#devtools::install_github("richfitz/datastorr")
#devtools::install_github("traitecoevo/baad.data")
```

## Review the data:



```{r data, echo=TRUE, eval=TRUE}
baad <- baad.data::baad_data()
d_baad <- baad$data
head(d_baad)
```

Refine (subset) the data for Pinus Ponderosa:

```{r subset_pipo, echo=TRUE, eval=TRUE}
pipo <- d_baad[ which(d_baad$species == 'Pinus ponderosa'), ]
#head(pipo)

```

Plot raw data, dbh to height.

```{r d2h_raw, echo=TRUE, eval=TRUE}
# Set variables from the data. 

# Diameter at breat height (dbh) will be used for most allometric calculations
dbh <- (pipo$d.bh)*100
# Height
h <- (pipo$h.t)

plot(h~dbh,
     main="Diameter at breast height (dbh) to Height",
     xlab = "dbh (mm)",
     ylab = "Height (m)")
```

FATES uses the dbh to height relationship from O'Brien et al., 1995. The log10Height (*m*) is regressed on log10DBH (*mm*). 
Where slope is p1 or fates_allom_d2h1 and intercept is p2  or fates_allom_d2h2 in the parameter file.

$$ log_{10}Height = log_{10}DBH*slope + intercept $$
So modeled height would be. I want to solve for the best fit p1 and p2 given height and dbh data from BAAD.

$$ Height = 10^{(log10(min(d,dbhmax))*p1 + p2)} $$

```{r simplemod_d2h, echo=TRUE, eval=TRUE}
# But first I will test out just a linear model
d2hmod <- lm(h~dbh)
coef(d2hmod)

# Then a linear model with log10(h) regressed on log10(dbh)
d2hmodlog <- lm(log10(h)~log10(dbh))
coef(d2hmodlog)
```

Next, I will create a function to represent the O'Brien calculation used for dbh to height relationships.

```{r obrien, echo=TRUE, eval=TRUE}
obrien <- function(dbh, p1, p2){
  height <- 10^((log10(dbh))* p1 + p2)
  return(height)
}

dbins <- c(10,20,30,40,50,60,70,80,90) # Remember in O'Brien dbh is in mm. 

# Use default parameters from O'Brien et al 1995
default <- obrien(dbins, 0.7, -0.2)
# Use the parameters from the log10 linear model d2hmodlog
lmfit <- obrien(dbins, 0.82, 0.004)

coef(lm(default~dbins))
coef(lm(lmfit~dbins))

plot(h~dbh,
     main="Diameter at breast height (dbh) to Height",
     xlab = "dbh (mm)",
     ylab = "Height (m)")

points(default~dbins, col="red", pch=16) # default O'Brien slope (0.7) and intercept (-0.2)
points(lmfit~dbins, col="blue", pch=16) # BAAD data slope (0.82) and intercept (-0.82)
#curve(1.89+0.05*x, add=T, col="blue",  lwd=1) # linear model coefficients 
#curve(-0.82 + (0.82*x), add=T, col="green", lwd=2) # log regression coefficients from d2hmodlog
#curve(10^(-0.2)+(10^0.7*x), add=T, col="red", lwd=4) # linear model using default coefficients from O'Brien et al 1995
#curve(11.7 + 0.07*x, add=T, col="red", lty=2, lwd=2) # linear model using default coefficients from O'Brien et al 1995
#curve(3.46 + 0.04*x, add=T, col="blue", lty=2, lwd=2) # linear fit of slope and intercept to points modeled from (d2hmodlog)

```

