---
title: "Pondo_Allom_Params"
author: "Katie Murenbeeld"
date: "3/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:

Information about allometric curves and functions used in FATES.
https://fates-docs.readthedocs.io/en/latest/fates_tech_note.html#allometry-and-growth-along-allometric-curves

## Install the BAAD Package:

Install the package from github:

```{r install, echo=TRUE, eval=TRUE}
#install.packages("devtools")
#devtools::install_github("richfitz/datastorr")
#devtools::install_github("traitecoevo/baad.data")
```

## Review the data:

```{r data, echo=TRUE, eval=TRUE}
baad <- baad.data::baad_data()
d_baad <- baad$data
head(d_baad)
```

Refine (subset) the data for Pinus Ponderosa:

```{r subset_pipo, echo=TRUE, eval=TRUE}
pipo <- d_baad[ which(d_baad$species == 'Pinus ponderosa'), ]
#head(pipo)

```

Plot raw data, dbh to height.

```{r d2h_raw, echo=TRUE, eval=TRUE}
# Set variables from the data. 

# Diameter at breat height (dbh) will be used for most allometric calculations
dbh_mm <- (pipo$d.bh)*1000
dbh_cm <- (pipo$d.bh)*100
# Height
h <- (pipo$h.t)

plot(h~dbh_mm,
     main="Diameter at breast height (dbh) to Height",
     xlab = "dbh (mm)",
     ylab = "Height (m)")

hist(dbh_mm)
hist(log10(dbh_mm))
```
\newpage

## Diameter at breast height (dbh) to height - d2h

FATES allows for four different approaches to predicting height from dbh:

* A power function $$ h = p_{1} * d^(p_{2}) $$
* O'Brien et al. (1995) $$ log_{10}Height = log_{10}DBH*slope + intercept $$
* Poorter et al. (2006) $$ h = p_{1} * (1-exp(p_{2}*d^(p_{3}))) $$
* Martinex Cano et al. (2019) $$ h = (p_{1}*d^(p_{2})/(p_{3} + d^(p_{2}))) $$


For my research I use the dbh to height relationship from O'Brien et al., 1995. The log10Height (*m*) is regressed on log10DBH (*mm*). 
Where slope is p1 or fates_allom_d2h1 and intercept is p2  or fates_allom_d2h2 in the parameter file.

$$ log_{10}Height = log_{10}DBH*slope + intercept $$
So modeled height would be. I want to solve for the best fit p1 and p2 given height and dbh data from BAAD.

$$ Height = 10^{(log10(min(d,dbhmax))*p1 + p2)} $$

```{r simplemod_d2h, echo=TRUE, eval=TRUE}
# But first I will test out just a linear model
d2hmod_mm <- lm(h~dbh_mm)
coef(d2hmod_mm)

# Then a linear model with log10(h) regressed on log10(dbh)
d2hmodlog_mm <- lm(log10(h)~log10(dbh_mm))
coef(d2hmodlog_mm)

d2hmodlog_cm <- lm(log10(h)~log10(dbh_cm))
coef(d2hmodlog_cm)
```

Next, I will create a function to represent the O'Brien calculation used for dbh to height relationships.

```{r obrien, echo=TRUE, eval=TRUE}
obrien <- function(dbh, p1, p2){
  height <- 10^((log10(dbh))* p1 + p2)
  return(height)
}

dbins_mm <- c(100,200,300,400,500,600,700,800,900) # Remember in O'Brien dbh is in mm. 
dbins_cm <- c(10,20,30,40,50,60,70,80,90) # But does FATES do that conversion or use cm?

# Use default parameters from O'Brien et al 1995
default_mm <- obrien(dbins_mm, 0.7, -0.2)
# Use the parameters from the log10 linear model d2hmodlog
lmfit_mm <- obrien(dbins_mm, 0.82, -0.82) # dbh in mm
lmfit_cm <- obrien(dbins_cm, 0.82, 0.003) # dbh in cm

coef(lm(default_mm~dbins_mm))
coef(lm(lmfit_mm~dbins_mm))
coef(lm(lmfit_cm~dbins_cm))

plot(h~dbh_mm,
     main="Diameter at breast height (dbh) to Height",
     xlab = "dbh (mm)",
     ylab = "Height (m)")

points(default_mm~dbins_mm, col="red", pch=16) # default O'Brien slope (0.7) and intercept (-0.2)
points(lmfit_mm~dbins_mm, col="blue", pch=16) # BAAD data slope (0.82) and intercept (-0.82)

```

```{r log10plot, echo=TRUE, eval=TRUE}

plot(log10(h)~log10(dbh_mm),
     main="Diameter at breast height (dbh) to Height",
     xlab = "log10 dbh (mm)",
     ylab = "log10 Height (m)")
curve(-0.82 + (0.82*x), add=T, col="blue", lwd=2) # log regression coefficients from d2hmodlog
```
```{r plots_d2h_cm, echo=TRUE, eval=TRUE}

plot(h~dbh_cm,
     main="Diameter at breast height (dbh) to Height",
     xlab = "dbh (cm)",
     ylab = "Height (m)")

points(default_mm~dbins_mm, col="red", pch=16) # default O'Brien slope (0.7) and intercept (-0.2)
points(lmfit_cm~dbins_cm, col="blue", pch=16) # BAAD data slope (0.82) and intercept (-0.82)

plot(log10(h)~log10(dbh_cm),
     main="Diameter at breast height (dbh) to Height",
     xlab = "log10 dbh (cm)",
     ylab = "log10 Height (m)")
curve(0.003 + (0.82*x), add=T, col="blue", lwd=2) # log regression coefficients from d2hmodlog

```



\newpage

## Diameter at breast height (dbh) to above ground biomass (AGB) - d2bagw

FATES provides three different options for calculating AGB from dbh:

* Saldarriaga et al. (1998) $$ (C_{agb} = f_{agb} * p_{1} * h^{p_{2}} * d^{p_{3}} * rho^{p_{4}}) $$
* 2 parameter power function $$ (C_{agb} = p_{1}/c2b * d^{p_{2}} )$$
* Chave et al. (2014) $$ (C_{agb} = p_{1}/c2b * (rho * d^2 * h)^{p_{2}})$$


For this project I will use Saldarriaga et al. (1998) 
$$ C_{agb} = f_{agb} * p_{1} * h^{p_{2}} * d^{p_{3}} * rho^{p_{4}}$$

```{r d2bagw, echo=TRUE, eval=TRUE}

# Define the variables needed for the Saldarriaga function
rho = 0.367
f_agb = 0.6
d.agb <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
h.agb <- c(2.82, 7.7, 13.9, 21.0, 29.1, 37.9, 37.9, 37.9, 37.9, 57.5)
#use heights from d2h
lmfit.h <- obrien(d.agb, 0.82, -0.82)
choj.h <- c(21.2, 117.1, 318, 646.1, 1119.6, 1754.5, 2565.1, 3564.4, 4764.4, 6176.6)

# Define the Saldarriaga function 
sal <- function(f_agb, p1, h, p2, d, p3, rho, p4){
  agb <- f_agb * p1 * h^p2 * d^p3 * rho^p4
  return(agb)
}

# Define a 2 parameter power function (as a stand-in for Chojnakcy if those param values are used)
par2_pwr <- function(p1,d,p2,c2b){
  bagw <- (p1*(d^p2))/c2b
  return(bagw)
}

choj_eq <- function(b0,b1,d){
  biom <- b0  + b1*log(d)
  return(biom)
}
choj_p1 = 2.6177
choj_p2 = 2.4638
c2b = 2

test_2par <- par2_pwr(choj_p1, d.agb, choj_p2, c2b)
test_choj <- choj_eq(choj_p1, choj_p2, d.agb)

# Parameters from earlier Jupyter Notebook. Not sure how to determine the best fit for the Sal function in r.
p1 = 0.131
p2 = 0.626
p3 = 2.46
p4 = 2.18

test <- sal(f_agb, p1, h.agb, p2, d.agb, p3, rho, p4)
test2 <- sal(f_agb, p1, lmfit.h, p2, d.agb, p3, rho, p4)
#plot(choj.h~d.agb, type="l", col="blue")
plot(test~d.agb, col="red", type="l", ylab="agb", xlab="dbh (cm)")
points(test2~d.agb, col="yellow", pch=16)
points(test_2par~d.agb, col="pink", pch=16)
points(test_choj~d.agb, col="green", pch=16)


```










